An asymptotically normal out-of-sample test based on mixed estimation windows

### 2016-01-13 mixedwindow.tex
Change url to GitHub repository
Why: the current version of the repository refers to a private
GitLab repository; it's more useful to direct people to the
public GitHub repository.

This change addresses this need by: changing the url to GitHub.

Closes #13
Add Table of Contents to the paper
Why: the working paper version of this document will be easier
for people to skim if it has a table of contents. This commit
adds one and also manually adds the assumptions and theoretical
results to the ToC.

\author{Gray Calhoun\thanks{Economics Department; Iowa State
    University; Ames, IA 50011.  Telephone: (515) 294-6271.  Email:
    \guillemotleft \protect\url{gcalhoun@iastate.edu}\guillemotright,
    web: \guillemotleft http://gray.clhn.org\guillemotright.
    If you find errors or mistakes in this paper, please email me or
    open a new issue at
    \guillemotleft https://git.ece.iastate.edu/gcalhoun/mixedwindow/issues\guillemotright.
    I'd like to
    thank Helle Bunzel, Todd Clark, Graham Elliott, Yu-Chin Hsu,
    Michael McCracken, Pablo Pincheira, Elie Tamer, Allan Timmermann, Ken West, Stephane
    Meng-Feng Yen, several referees, and participants at the 2011 Midwest Econometrics
    Group meeting and the 2013 \nber-\nsf\ Time Series conference
    for helpful comments and discussions.  I'd also like to thank Amit
    Goyal for providing computer code and data for his 2008
    RFS paper with Ivo Welch \citep{GoW:08}. An earlier version of this paper
    circulated under the name ``An asymptotically normal out-of-sample
  test of equal predictive accuracy for nested models.''}\\%
  Iowa State University}

@@

\begin{abstract}
  \noindent This paper develops a modification of \citepos[\textit{J.
    Econom.}]{ClW:07} adjusted out-of-sample $t$-test. We propose
  using a recursive window to estimate the benchmark model but a
  fixed-length rolling window to estimate the alternative. The
  resulting statistic is asymptotically normal even when the models
  are nested. The
  paper also presents Monte Carlo evidence that this statistic has
  much higher power than existing out-of-sample statistics in a common
  use-case for these tests: when the \dgp\ is subject to instability.
  This procedure is then used to analyze
  \citepos[\textit{Rev. Finan. Stud.}]{GoW:08} excess returns dataset
  and supports their finding that the equity premium is unpredictable
  out-of-sample.

@@

\end{abstract}

@@

\begin{asmp}\label{a1}%
  The data are generated by the relationship
  \begin{equation}
    y_{t+1} = x_t'\btrue + \ep_{t+1}
  \end{equation}
  for $t=1,2,\dots$, for some value $\btrue$, with $\E x_t \ep_{t+1} =
  0$, $\E \ep_{t+1}^2 > 0$, and $\E x_t x_t'$ positive definite for
  all $t$. Also assume that there is an additional sequence of random
  vectors $z_t$ and the process $(\ep_{t+1}, x_t, z_t)$ is stationary
  and strong mixing of size $-r/(r-2)$ or uniform mixing of size
  $-r/(2r-2)$, for $r > 2$.
\end{asmp}

@@

\begin{asmp}\label{a3}%
  The benchmark forecast is $x_t'\bh_t$, where $\bh_t$ is constructed
  with a recursive window according to~\eqref{eq:8}. The alternative
  forecast satisfies
  \begin{equation}
    \yh_{t+1} = \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})
  \end{equation}
  where $\psi$ is a known measurable function and the window length,
  $R$, remains finite as $T \to \infty$. Moreover, the vector
  $(\ep_{t+1}, x_t, \yh_{t+1})$ has uniformly bounded $2 r$ moments
  where $r$ is first defined in Assumption~\ref{a1}.
\end{asmp}

@@

Our next assumption ensures that the asymptotic variance of the \oos\
average is positive.
\begin{asmp}\label{a4}%
  The asymptotic variance-covariance matrix
  \begin{equation}
    \var \Bigg(
      \oclt{t} \begin{pmatrix} x_t \\ \yh_{t+1} \end{pmatrix} \ep_{t+1}
      \Bigg)
  \end{equation}
  is uniformly positive definite (in $T$).
\end{asmp}
This assumption is much less restrictive than in \cite{Wes:96}.  As in
\cite{GiW:06} and \citet{ClW:06,ClW:07}, the assumption only serves to
rule out pathological cases --- for example, letting the alternative
model consist of only the first regressor of the benchmark. In \citet{Wes:96}, this
assumption is a restriction on the \dgp\ as well as the forecasting
models, but in this paper it is a restriction only on the models.

The final assumption restricts the class of \hac\ variance estimators
we will consider. We use the same class of estimators studied by
\citet{JoD:00} (their class $\mathcal{K}$); see their paper for
further discussion.
\begin{asmp}\label{a5}%
  The kernel $K$ is a function from $\Re$ to $[-1,1]$ such that $K(0) = 1$, $K(x)
  = K(-x)$ for all $x$, $K(\cdot)$ is continuous at zero and all but a
  finite number of points, and
  \begin{gather*}
    \int_{-\infty}^{\infty} \lvert K(x) \rvert\, dx < \infty,
    \intertext{and}
    \int_{-\infty}^{\infty} \Bigg\lvert
    \int_{-\infty}^{\infty} K(z) e^{ixz}\,dz \Bigg\rvert\, dx < \infty.
  \end{gather*}
\end{asmp}

@@

\begin{thm}\label{res:1}\input{mixedwindow_thm1}\end{thm}

@@

\begin{lem}\label{lem:2}\input{mixedwindow_lem2}\end{lem}

@@

\begin{thm}\label{thm:3}\input{mixedwindow_thm3}\end{thm}

@@

\appendix
\section{Not-for-publication supplemental appendix}
\newcommand{\WesA}[1][]{\ocltb{t}
  (F_t^{#1} - F) B^{#1} H_t^{#1}}
\newcommand{\WesB}[1][]{\ocltb{t} F (B_t^{#1} -
  B^{#1}) H_t^{#1}}
\newcommand{\WesC}[1][]{\ocltb{t}
  (F_t^{#1} - F) (B_t^{#1} - B^{#1}) H_t^{#1}}

@@

\subsection*{Proof of Theorem \ref{res:1}}
  Let $R'$ be a new sequence such that $R' \to \infty$ as $T \to \infty$
  and $R' = o(\sqrt{P})$, and then rewrite the centered \oos\ average as
  \begin{equation}\label{eq:6}
    \sqrt{P} (\fb - \E \fb^*)
    = \ocltb{t} ((f_t - \E f_t) + (\fh_t - f_t))
      + \tfrac{1}{\sqrt{P}} \osumc{t} (\fh_t - \E f_t).
  \end{equation}
  Lemma~\ref{res:a1} ensures that the second summation is $o_p(1)$, so
  we can use a Taylor expansion to rewrite~\eqref{eq:6} as
  \begin{align*}
    \sqrt{P} (\fb - \E \fb^*)
    &= \ocltb{t} (f_t - \E f_t) + F B \ocltb{t} H_t \\
    & \quad + \WesA + \WesB \\ & \quad + \WesC + \oclt{t} w_t + o_p(1)
  \end{align*}
  where $w_t$ equals $2 (\bh_t - \btrue)' x_t x_t' (\bh_t - \btrue)$.
  Lemma~\ref{res:a4} shows that
  \begin{gather}
    \WesA \to^{p} 0 \label{eq:11} \\
    \WesB \to^{p} 0 \label{eq:12} \\
    \intertext{and}
    \WesC \to^{p} 0 \label{eq:13}
  \end{gather}
  and Lemma~\ref{res:a2} along with the \clt\ ensures that $\oclt{t}
  w_t = o_{p}(1)$. The proof that
  \begin{equation*}
    \ocltb{t} (f_t - \E f_t) + F B \ocltb{t} H_t \to N(0, \sigma^2).
  \end{equation*}
  follows the same argument as in \citet{Wes:96} and \citet{Mcc:00}.

\subsection*{Proof of Lemma~\ref{lem:2}}

@@

\subsection*{Supporting Results}

\begin{alem}\label{res:a1}
  Suppose the conditions of Theorem~\ref{res:1} hold, and define $R'$
  to be a sequence that satisfies $R' \to \infty$ as $T \to \infty$
  and $R' = o(\sqrt{P})$. Then
  \begin{equation*}
    \ocltc{t} (\fh_t - \E f_t) \to^p 0.
  \end{equation*}
\end{alem}

@@

\begin{alem}\label{res:a2}
  Suppose $a \in [0,1/2)$ and Assumptions~\ref{a1}~--~\ref{a4}
  hold, and let $R'$ be a sequence such that $R' \to \infty$ as $T \to
  \infty$ and $R' = o(\sqrt{P})$. Then
  \begin{enumerate}
  \item $\omaxb{t} | (t-1)^a H_t | \to^p 0$,
  \item $\omaxc{t} | (t-1)^a H_t | = O_p(1)$,
  \item $\omaxb{t} | B_t - B | \to^p 0$,
  \item $\omaxc{t} | B_t - B | = O_p(1)$,
  \item $\omaxb{t} | (t-1)^a(\bh_t - \btrue) | \to^{p} 0$, and
  \item $\omaxc{t} | (t-1)^a(\bh_t - \btrue) | = O_p(1)$,
  \end{enumerate}
  where the absolute value is taken as the largest of the
  element-by-element absolute values.
\end{alem}

@@

\begin{alem}\label{res:a4}
  Under the conditions of Theorem~\ref{res:1}, Equations
  \eqref{eq:11}--\eqref{eq:13} hold.
\end{alem}

>>>>

\providecommand\phantomsection{}

\author{Gray Calhoun\thanks{Economics Department; Iowa State
    University; Ames, IA 50011.  Telephone: (515) 294-6271.  Email:
    \guillemotleft \protect\url{gcalhoun@iastate.edu}\guillemotright,
    web: \guillemotleft http://gray.clhn.org\guillemotright.
    If you find errors or mistakes in this paper, please email me or
    open a new issue at
    \guillemotleft https://github.com/grayclhn/mixedwindow/issues\guillemotright.
    I'd like to
    thank Helle Bunzel, Todd Clark, Graham Elliott, Yu-Chin Hsu,
    Michael McCracken, Pablo Pincheira, Elie Tamer, Allan Timmermann, Ken West, Stephane
    Meng-Feng Yen, several referees, and participants at the 2011 Midwest Econometrics
    Group meeting and the 2013 \nber-\nsf\ Time Series conference
    for helpful comments and discussions.  I'd also like to thank Amit
    Goyal for providing computer code and data for his 2008
    RFS paper with Ivo Welch \citep{GoW:08}. An earlier version of this paper
    circulated under the name ``An asymptotically normal out-of-sample
  test of equal predictive accuracy for nested models.''}\\%
  Iowa State University}

@@

\addcontentsline{toc}{section}{Abstract}
\begin{abstract}
  \noindent This paper develops a modification of \citepos[\textit{J.
    Econom.}]{ClW:07} adjusted out-of-sample $t$-test. We propose
  using a recursive window to estimate the benchmark model but a
  fixed-length rolling window to estimate the alternative. The
  resulting statistic is asymptotically normal even when the models
  are nested. The
  paper also presents Monte Carlo evidence that this statistic has
  much higher power than existing out-of-sample statistics in a common
  use-case for these tests: when the \dgp\ is subject to instability.
  This procedure is then used to analyze
  \citepos[\textit{Rev. Finan. Stud.}]{GoW:08} excess returns dataset
  and supports their finding that the equity premium is unpredictable
  out-of-sample.

@@

\end{abstract}
\newpage

\tableofcontents

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Assumption \ref{a1}}
\begin{asmp}\label{a1}%
  The data are generated by the relationship
  \begin{equation}
    y_{t+1} = x_t'\btrue + \ep_{t+1}
  \end{equation}
  for $t=1,2,\dots$, for some value $\btrue$, with $\E x_t \ep_{t+1} =
  0$, $\E \ep_{t+1}^2 > 0$, and $\E x_t x_t'$ positive definite for
  all $t$. Also assume that there is an additional sequence of random
  vectors $z_t$ and the process $(\ep_{t+1}, x_t, z_t)$ is stationary
  and strong mixing of size $-r/(r-2)$ or uniform mixing of size
  $-r/(2r-2)$, for $r > 2$.
\end{asmp}

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Assumption \ref{a3}}
\begin{asmp}\label{a3}%
  The benchmark forecast is $x_t'\bh_t$, where $\bh_t$ is constructed
  with a recursive window according to~\eqref{eq:8}. The alternative
  forecast satisfies
  \begin{equation}
    \yh_{t+1} = \psi(y_t,z_t,\dots,y_{t-R+1}, z_{t-R+1})
  \end{equation}
  where $\psi$ is a known measurable function and the window length,
  $R$, remains finite as $T \to \infty$. Moreover, the vector
  $(\ep_{t+1}, x_t, \yh_{t+1})$ has uniformly bounded $2 r$ moments
  where $r$ is first defined in Assumption~\ref{a1}.
\end{asmp}

@@

Our next assumption ensures that the asymptotic variance of the \oos\
average is positive.
\phantomsection
\addcontentsline{toc}{subsubsection}{Assumption \ref{a4}}
\begin{asmp}\label{a4}%
  The asymptotic variance-covariance matrix
  \begin{equation}
    \var \Bigg(
      \oclt{t} \begin{pmatrix} x_t \\ \yh_{t+1} \end{pmatrix} \ep_{t+1}
      \Bigg)
  \end{equation}
  is uniformly positive definite (in $T$).
\end{asmp}
This assumption is much less restrictive than in \cite{Wes:96}.  As in
\cite{GiW:06} and \citet{ClW:06,ClW:07}, the assumption only serves to
rule out pathological cases --- for example, letting the alternative
model consist of only the first regressor of the benchmark. In \citet{Wes:96}, this
assumption is a restriction on the \dgp\ as well as the forecasting
models, but in this paper it is a restriction only on the models.

The final assumption restricts the class of \hac\ variance estimators
we will consider. We use the same class of estimators studied by
\citet{JoD:00} (their class $\mathcal{K}$); see their paper for
further discussion.
\phantomsection
\addcontentsline{toc}{subsubsection}{Assumption \ref{a5}}
\begin{asmp}\label{a5}%
  The kernel $K$ is a function from $\Re$ to $[-1,1]$ such that $K(0) = 1$, $K(x)
  = K(-x)$ for all $x$, $K(\cdot)$ is continuous at zero and all but a
  finite number of points, and
  \begin{gather*}
    \int_{-\infty}^{\infty} \lvert K(x) \rvert\, dx < \infty,
    \intertext{and}
    \int_{-\infty}^{\infty} \Bigg\lvert
    \int_{-\infty}^{\infty} K(z) e^{ixz}\,dz \Bigg\rvert\, dx < \infty.
  \end{gather*}
\end{asmp}

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Theorem \ref{res:1}}
\begin{thm}\label{res:1}\input{mixedwindow_thm1}\end{thm}

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Lemma \ref{lem:2}}
\begin{lem}\label{lem:2}\input{mixedwindow_lem2}\end{lem}

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Theorem \ref{thm:3}}
\begin{thm}\label{thm:3}\input{mixedwindow_thm3}\end{thm}

@@

\newpage
\phantomsection
\addcontentsline{toc}{section}{References}
\bibliography{texextra/references}

\newpage
\appendix
\section{Supplemental appendix (not for publication)}
\newcommand{\WesA}[1][]{\ocltb{t}
  (F_t^{#1} - F) B^{#1} H_t^{#1}}
\newcommand{\WesB}[1][]{\ocltb{t} F (B_t^{#1} -
  B^{#1}) H_t^{#1}}
\newcommand{\WesC}[1][]{\ocltb{t}
  (F_t^{#1} - F) (B_t^{#1} - B^{#1}) H_t^{#1}}

@@

\subsection{Proof of Theorem \ref{res:1}}
  Let $R'$ be a new sequence such that $R' \to \infty$ as $T \to \infty$
  and $R' = o(\sqrt{P})$, and then rewrite the centered \oos\ average as
  \begin{equation}\label{eq:6}
    \sqrt{P} (\fb - \E \fb^*)
    = \ocltb{t} ((f_t - \E f_t) + (\fh_t - f_t))
      + \tfrac{1}{\sqrt{P}} \osumc{t} (\fh_t - \E f_t).
  \end{equation}
  Lemma~\ref{res:a1} ensures that the second summation is $o_p(1)$, so
  we can use a Taylor expansion to rewrite~\eqref{eq:6} as
  \begin{align*}
    \sqrt{P} (\fb - \E \fb^*)
    &= \ocltb{t} (f_t - \E f_t) + F B \ocltb{t} H_t \\
    & \quad + \WesA + \WesB \\ & \quad + \WesC + \oclt{t} w_t + o_p(1)
  \end{align*}
  where $w_t$ equals $2 (\bh_t - \btrue)' x_t x_t' (\bh_t - \btrue)$.
  Lemma~\ref{res:a4} shows that
  \begin{gather}
    \WesA \to^{p} 0 \label{eq:11} \\
    \WesB \to^{p} 0 \label{eq:12} \\
    \intertext{and}
    \WesC \to^{p} 0 \label{eq:13}
  \end{gather}
  and Lemma~\ref{res:a2} along with the \clt\ ensures that $\oclt{t}
  w_t = o_{p}(1)$. The proof that
  \begin{equation*}
    \ocltb{t} (f_t - \E f_t) + F B \ocltb{t} H_t \to N(0, \sigma^2).
  \end{equation*}
  follows the same argument as in \citet{Wes:96} and \citet{Mcc:00}.

\subsection{Proof of Lemma~\ref{lem:2}}

@@

\subsection{Supporting results}

\phantomsection
\addcontentsline{toc}{subsubsection}{Lemma \ref{res:a1}}
\begin{alem}\label{res:a1}
  Suppose the conditions of Theorem~\ref{res:1} hold, and define $R'$
  to be a sequence that satisfies $R' \to \infty$ as $T \to \infty$
  and $R' = o(\sqrt{P})$. Then
  \begin{equation*}
    \ocltc{t} (\fh_t - \E f_t) \to^p 0.
  \end{equation*}
\end{alem}

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Lemma \ref{res:a2}}
\begin{alem}\label{res:a2}
  Suppose $a \in [0,1/2)$ and Assumptions~\ref{a1}~--~\ref{a4}
  hold, and let $R'$ be a sequence such that $R' \to \infty$ as $T \to
  \infty$ and $R' = o(\sqrt{P})$. Then
  \begin{enumerate}
  \item $\omaxb{t} | (t-1)^a H_t | \to^p 0$,
  \item $\omaxc{t} | (t-1)^a H_t | = O_p(1)$,
  \item $\omaxb{t} | B_t - B | \to^p 0$,
  \item $\omaxc{t} | B_t - B | = O_p(1)$,
  \item $\omaxb{t} | (t-1)^a(\bh_t - \btrue) | \to^{p} 0$, and
  \item $\omaxc{t} | (t-1)^a(\bh_t - \btrue) | = O_p(1)$,
  \end{enumerate}
  where the absolute value is taken as the largest of the
  element-by-element absolute values.
\end{alem}

@@

\phantomsection
\addcontentsline{toc}{subsubsection}{Lemma \ref{res:a4}}
\begin{alem}\label{res:a4}
  Under the conditions of Theorem~\ref{res:1}, Equations
  \eqref{eq:11}--\eqref{eq:13} hold.
\end{alem}
